---
title: "Edinburgh Biodiversity"
author: Emma Gemal
date: 24/09/2019
output: html_document 
---

```{r,include=FALSE} 
library(rmarkdown)   #loading the packages necessary for this exercise
```
```{r,include=FALSE}
library(dplyr)
```

```{r, include=FALSE}
setwd("//Users/emmagemal/Documents/Uni Work/Year 3/Data Science/CC-2-RMarkdown-master")   #working directory set
edidiv<-read.csv("edidiv.csv")
```



*The data used in this report is a publicly available dataset of occurrence records for many animal, plant, and fungi species, for 2000-2016 from the NBN Gateway.*

### Species richness in Edinburgh, Scotland is displayed below. 
The `richness` was calculated by:

```{r}
richness <- 
  edidiv %>%
    group_by(taxonGroup) %>%
    summarise(Species_richness = n_distinct(taxonName))
```
The results for `richness` are displayed in Table 1.

#### Table 1.
```{r,echo=FALSE}
library(pander)    #above is the construction of a table of species richness for each taxanomic group
pander(richness)   #table displayed nicely via pander 
```

This is graphically displayed below:

```{r,echo=FALSE}
barplot(richness$Species_richness, 
        names.arg = richness$taxonGroup, 
        xlab = "Taxa", ylab = "Number of species", 
        ylim = c(0,600)
        )                       #creating a barchart of taxa and species abundance  
```

The most common species (`max_abund`) for each taxanomic group was also found, displayed in Table 2. The code used for this is shown here:
```{r}
max_abund <-
  edidiv %>%
    group_by(taxonGroup) %>%
    summarise(taxonName = names(which.max(table(taxonName))))
```
#### Table 2.
```{r,echo=FALSE}
emphasize.italics.cols(2)    #making the 2nd column italics 
pander(max_abund)
```

The two parameters, `richness` and `max_abund`, were joined together to produce Table 3:
```{r}
richness_abund <- inner_join(richness, max_abund, by = "taxonGroup")
```
#### Table 3. 
```{r,echo=FALSE}
richness_abund <- rename(richness_abund, Most_abundant =  taxonName, Taxon = taxonGroup)
emphasize.italics.cols(3)
pander(richness_abund)
```

